Human Factors Influencing Visual Statistical Inference

### 2017-10-19 sociology_paper.Rnw
results updated

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Sun Aug 30 22:43:05 2015
\begin{table}[ht]
\caption{\label{tbl:anova_factor}Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable. Gender does not have a significant effect on detection rate, but does on time to respond. All factors significantly affect time to respond.}
\centering
\scalebox{0.9}{
\begin{tabular}{lrrrrlrrrr}
& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
\cline{2-5} \cline{7-10} 
 & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
Full  & 51872.4 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 23792.0 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
 - Age & 51329.3 & 543.2 & 6 & $<$0.001 & \hspace{0.5cm} & 23766.4 & 26.0 & 6 & $<$0.001 \\ 
 - Country & 51496.4 & 376.1 & 2 & $<$0.001 &  & 23765.9 & 26.0 & 2 & $<$0.001 \\ 
 - Degree & 51756.2 & 116.2 & 4 & $<$0.001 &  & 23768.2 & 23.8 & 4 & $<$0.001 \\ 
 - Gender & 51804.3 & 68.1 & 1 & $<$0.001 &  & 23789.7 & 2.2 & 1 & 0.140 \\ 
\cline{2-5} \cline{7-10}
\end{tabular}}
\end{table}

Parameter estimates from the model fits are shown in Table~\ref{tbl:model_result_demographics}. The first level of each factor serves as the baseline for the model. The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for age groups 31-35 and above 50. Participants from India have a similar detection rate to those from the USA, but participants from the rest of the world show a significantly higher detection rate. Subjects  reporting to have  a graduate degree had a significantly higher detection rate. For time to respond, all age groups were significantly slower than  18-25 year olds.  Participants from India were slower, while those from the rest of the world were faster than US Americans. A higher degree corresponded to longer time, and males were on average slower than females.

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Sat Aug 29 14:17:22 2015
\begin{table}[hbtp]
\caption{\label{tbl:model_result_demographics}Parameter estimates of models~\eqref{eqn:demographic_time} and~\eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively. For time taken all the demographic factors are significant. For detection rate, age group 36-40, rest of the world and graduate degree are significantly different. For gender no difference in performance is observed. Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
\centering
\scalebox{.9}{
\begin{tabular}{rrr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
 && \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\

\cline{3-6} \cline{8-11} 
\multicolumn{2}{c}{{Demographic Factor} \hfill Level} & Est & 2.5\% & 97.5\% &   && Est & 2.5\% & 97.5\% &   \\ 
\cline{3-6} \cline{8-11}
\bf Fixed Effects & Average $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.61 & -0.79 & -0.43 & *** \\ [3pt]
\bf Age Category $(\alpha)$ & 18-25 & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
&  26-30 & 0.06 & 0.03 & 0.08 & *** &   & 0.06 & -0.03 & 0.16 &  \\ 
&  31-35 & 0.07 & 0.04 & 0.10 & *** &   & 0.11 & 0.01 & 0.22 & * \\ 
&  36-40 & 0.23 & 0.20 & 0.26 & *** &   & 0.31 & 0.19 & 0.43 & *** \\ 
&  41-45 & 0.18 & 0.14 & 0.22 & *** &   & 0.16 & 0.00 & 0.31 & * \\ 
&  46-50 & 0.27 & 0.22 & 0.32 & *** &   & 0.14 & -0.05 & 0.33 &  \\ 
&  above 50 & 0.35 & 0.32 & 0.39 & *** &   & 0.15 & 0.01 & 0.28 & * \\ [3pt]
\bf Country $(\kappa)$& U.S. & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
&  India & 0.23 & 0.21 & 0.25 & *** &   & -0.18 & -0.27 & -0.10 & *** \\ 
&  Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.08 & -0.04 & 0.20 &  \\ [3pt]
\bf Education $(\tau)$& High School & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
&  Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & -0.03 & -0.18 & 0.11 &  \\ 
&  Under grad degree & 0.13 & 0.10 & 0.17 & *** &   & 0.12 & -0.01 & 0.26 & . \\ 
&  Graduate courses & 0.01 & -0.04 & 0.05 &  &   & -0.17 & -0.33 & -0.01 & * \\ 
&  Graduate degree & 0.13 & 0.09 & 0.16 & *** &   & -0.06 & -0.19 & 0.07 &  \\ [3pt]
\bf Gender $(\gamma)$& Female & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
&  Male & 0.08 & 0.06 & 0.10 & *** &   & 0.06 & -0.02 & 0.13 &  \\
\cline{3-6} \cline{8-11}
\bf Random Effects &  lineup $(\sigma_\ell)$& \multicolumn{5}{l}{0.29} & \multicolumn{4}{l}{\phantom{-}2.29} \\ 
                   &  Error $(\sigma)$& \multicolumn{5}{l}{0.69} & \multicolumn{4}{l}{\phantom{-}0.89}  \\ 
\cline{3-6} \cline{8-11}\\
\hline
\multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
\end{tabular}}
\end{table}

@@

While some of the demographic factors are strongly statistically significant, the main source of variation in detection rate is the lineup difficulty. For example, let's examine the effect of graduate degree. To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with some graduate course work as compared to an undergraduate degree, for an average difficulty lineup (random effect of zero). Plugging in the relevant quantities to the fitted model gives a difference equal to:
$$ \frac {\exp(-0.61-0.17)}{1+\exp(-0.61-0.17)}-  \frac {\exp(-0.61+0.12)}{1+\exp(-0.61+0.12)} = 0.314- 0.379 =-0.065.$$ 
The person with some graduate course work picks the data plot  on average in 33.5\% of lineups of average difficulty, as compared to 37.9\% if they have an undergraduate degree. This difference is reduced to 2\% for a lineup with one standard deviation order of magnitude difference in difficulty. For two standard deviations it further reduces to 0.3\%. Thus, although there is a statistically significant difference in the proportion at which participants identify the  data plot for some demographic factors, these are not  significant in any practical sense.  Figure~\ref{fig:practical_impact_graduate} illustrates this example showing fitted models for a US 18-25 female with either a high school education or a graduate degree.  Similar calculations show the same negligible impact of age level 36-40 (\Sexpr{round(exp(-0.61+0.31)/(1+ exp(-0.61+0.31)) - exp(-0.61)/(1 + exp(-0.61)),3)} at most)  and country (\Sexpr{round(exp(-0.61+0.08)/(1+ exp(-0.61+0.08)) - exp(-0.61-0.18)/(1 + exp(-0.61-0.18)), 3)} at most) on the probability of a correct response. Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.

\begin{figure}[htbp] 
   \centering
%   \includegraphics[width=6.5in]{learning_trend.pdf} 
    \includegraphics[width=3.2in]{practical_impact_graduate.pdf} 
    \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with a graduate degree compared to a high school degree. Even though having a graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.045 which is  negligible from a practical point of view. The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} =2.29$) 
   of lineup variability.}
   \label{fig:practical_impact_graduate}
\end{figure}

@@

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Sun Aug 30 11:15:41 2015
\begin{table}[hbtp]
\centering
\caption{\label{tbl:model_result_response}Parameter estimates of models \eqref{eqn:trend_response} fitted to data from three different experiments using detection rate as the response to assess learning trend. Attempt number is fitted as a factor to enable modeling any sort of non-linear learning trend. Only experiment~6 shows some evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
\scalebox{.9}{
\begin{tabular}{rr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
\cline{2-5} \cline{7-10} \cline{12-15} 
& \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
\cline{2-5} \cline{7-10} \cline{12-15} 
Effect & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  \\ 
\cline{2-5} \cline{7-10} \cline{12-15} 
Fixed \\
$\mu$ & -1.30 & -1.69 & -0.92 & *** &   & -0.22 & -0.51 & 0.07 &  &   & -1.71 & -2.77 & -0.65 & ** \\ [3pt]
  $\alpha_2$ & 0.27 & -0.19 & 0.73 &  &   & 0.26 & -0.06 & 0.58 &  &   & -0.49 & -1.31 & 0.32 &  \\ 
  $\alpha_3$ & -0.18 & -0.65 & 0.29 &  &   & 0.34 & 0.02 & 0.66 & * &   & -0.19 & -1.01 & 0.63 &  \\ 
  $\alpha_4$ & 0.08 & -0.39 & 0.55 &  &   & 0.36 & 0.04 & 0.68 & * &   & -0.47 & -1.28 & 0.35 &  \\ 
  $\alpha_5$ & 0.30 & -0.17 & 0.77 &  &   & 0.37 & 0.05 & 0.70 & * &   & -0.16 & -0.96 & 0.65 &  \\ 
  $\alpha_6$ & 0.04 & -0.44 & 0.53 &  &   & 0.24 & -0.07 & 0.56 &  &   & -0.03 & -0.87 & 0.82 &  \\ 
  $\alpha_7$ & 0.28 & -0.20 & 0.77 &  &   & 0.16 & -0.16 & 0.48 &  &   & 0.02 & -0.81 & 0.84 &  \\ 
  $\alpha_8$ & -0.04 & -0.54 & 0.45 &  &   & 0.34 & 0.02 & 0.66 & * &   & -0.08 & -0.89 & 0.73 &  \\ 
  $\alpha_9$ & -0.20 & -0.69 & 0.30 &  &   & 0.38 & 0.05 & 0.70 & * &   & 0.15 & -0.71 & 1.02 &  \\ 
  $\alpha_{10}$ & 0.51 & 0.03 & 1.00 & * &   & 0.19 & -0.14 & 0.52 &  &   & -0.28 & -1.14 & 0.58 &  \\ 
  Random \\
  $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.85}  & \multicolumn{5}{l}{\phantom{-}0.90} & 0.76   \\ 
  $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.02}  & \multicolumn{5}{l}{\phantom{-}0.03} & 0.06   \\ 
  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}1.48}  & \multicolumn{5}{l}{\phantom{-}1.42} & 3.37   \\ \cline{2-5} \cline{7-10} \cline{12-15} 
\\
\hline
\multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
\end{tabular}}
\end{table}

@@

% latex table generated in R 3.2.2 by xtable 1.7-4 package
% Sun Aug 30 11:42:08 2015
\begin{table}[bhtp]
\caption{\label{tbl:model_result_time}Parameter estimates of model~\eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup. Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments~5,~6 and~7.}
\centering
\scalebox{0.9}{
\begin{tabular}{lr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
\cline{2-5} \cline{7-10} \cline{12-15} 
& \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\

>>>>

% latex table generated in R 3.4.1 by xtable 1.8-2 package
% Wed Oct 18 17:43:02 2017
\begin{table}[ht]
\caption{\label{tbl:anova_factor}Analysis of variance (ANOVA) table comparing full model, all the demographic factors, with the reduced models, obtained by removing respective factor variable. Gender does not have a significant effect on detection rate, but does on time to respond. All factors significantly affect time to respond.}
\centering
\scalebox{0.9}{
\begin{tabular}{lrrrrlrrrr}
& \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\
\cline{2-5} \cline{7-10} 
 & Deviance & $\chi^2$ & df & $p$-value && Deviance & $\chi^2$ & df & $p$-value \\ \cline{2-5} \cline{7-10}
Full  & 52609.1 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} &  & 24140.5 & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} & \multicolumn{1}{c}{---} \\ 
 - Age & 52081.3 & 527.8 & 6 & $<$0.001 & \hspace{0.5cm} & 24113.2 & 27.3 & 6 & $<$0.001 \\ 
 - Country & 52245.5 & 363.6 & 2 & $<$0.001 &  & 24114.8 & 25.7 & 2 & $<$0.001 \\ 
 - Degree & 52498.7 & 110.4 & 4 & $<$0.001 &  & 24116.0 & 24.5 & 4 & $<$0.001 \\ 
 - Gender & 52540.3 & 68.8 & 1 & $<$0.001 &  & 24138.9 & 1.6 & 1 & 0.200 \\ 
\cline{2-5} \cline{7-10}
\end{tabular}}
\end{table}

% % latex table generated in R 3.4.1 by xtable 1.8-2 package
% % Wed Oct 18 17:43:02 2017
% \begin{table}[ht]
% \centering
% \begin{tabular}{rrrrrlrrrr}
%   \hline
%  & deviance & Chisq & Chi.Df & Pr..Chisq. & g1 & deviance & Chisq & Chi.Df & Pr..Chisq. \\ 
%   \hline
% Full & 52609.06 & 0.00 & 6.00 & 0.00 &  & 24140.49 & 0.00 & 6.00 & 0.00 \\ 
%   Age & 52081.29 & 527.77 & 6.00 & 0.00 &  & 24113.24 & 27.25 & 6.00 & 0.00 \\ 
%   Country & 52245.46 & 363.60 & 2.00 & 0.00 &  & 24114.79 & 25.70 & 2.00 & 0.00 \\ 
%   Degree & 52498.65 & 110.41 & 4.00 & 0.00 &  & 24116.02 & 24.47 & 4.00 & 0.00 \\ 
%   Gender & 52540.30 & 68.76 & 1.00 & 0.00 &  & 24138.85 & 1.64 & 1.00 & 0.20 \\ 
%    \hline
% \end{tabular}
% \end{table}

Parameter estimates from the model fits are shown in Table~\ref{tbl:model_result_demographics}. The first level of each factor serves as the baseline for the model. The age group 36-40  has a significantly higher detection rate than the 18-25 age group, and to a lesser extent this is also true for other age groups except for 26-30. Participants from the rest of the world have a similar detection rate to those from the USA, but participants from India show a significantly lower detection rate. Subjects reporting to have  an undergraduate degree had a significantly higher detection rate. For time to respond, all age groups were significantly faster than  18-25 year olds.  Participants from India as well as the rest of the world were faster than US Americans. A higher degree corresponded to longer time, and males were on average faster than females.

% latex table generated in R 3.4.1 by xtable 1.8-2 package
% Wed Oct 18 17:22:24 2017
\begin{table}[hbtp]
\caption{\label{tbl:model_result_demographics}Parameter estimates of models~\eqref{eqn:demographic_time} and~\eqref{eqn:demographic_response} fitted for average log time taken and detection rate, respectively. For time taken all the demographic factors are significant. For detection rate, age group 36-40, rest of the world and graduate degree are significantly different. For gender no difference in performance is observed. Lineup variability is estimated to be very large for model~\eqref{eqn:demographic_response}.}
\centering
\scalebox{.9}{
\begin{tabular}{rrr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
 && \multicolumn{4}{c} {Log Time (model~\ref{eqn:demographic_time})} & &\multicolumn{4}{c} {Detection rate (model~\ref{eqn:demographic_response})}  \\

\cline{3-6} \cline{8-11} 
\multicolumn{2}{c}{{Demographic Factor} \hfill Level} & Est & 2.5\% & 97.5\% &   && Est & 2.5\% & 97.5\% &   \\ 
\cline{3-6} \cline{8-11}
\bf Fixed Effects & Average $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.63 & -0.81 & -0.45 & *** \\ [3pt]
\bf Age Category $(\alpha)$ & 18-25 & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
&  26-30 & 0.05 & 0.03 & 0.07 & *** &   & 0.03 & -0.06 & 0.13 &  \\ 
&  31-35 & 0.06 & 0.03 & 0.09 & *** &   & 0.12 & 0.02 & 0.23 & * \\ 
&  36-40 & 0.22 & 0.19 & 0.25 & *** &   & 0.29 & 0.17 & 0.41 & *** \\ 
&  41-45 & 0.17 & 0.13 & 0.21 & *** &   & 0.19 & 0.03 & 0.34 & * \\ 
&  46-50 & 0.28 & 0.23 & 0.32 & *** &   & 0.20 & 0.01 & 0.39 & * \\ 
&  above 50 & 0.34 & 0.31 & 0.38 & *** &   & 0.16 & 0.02 & 0.29 & * \\ [3pt]
\bf Country $(\kappa)$& U.S. & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
&  India & 0.23 & 0.20 & 0.25 & *** &   & -0.19 & -0.28 & -0.10 & *** \\ 
&  Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.06 & -0.05 & 0.18 &  \\ [3pt]
\bf Education $(\tau)$& High School & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
&  Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & 0.00 & -0.14 & 0.14 &  \\ 
&  Under grad degree & 0.14 & 0.10 & 0.17 & *** &   & 0.16 & 0.03 & 0.29 & * \\ 
&  Graduate courses & 0.02 & -0.02 & 0.06 &  &   & -0.13 & -0.29 & 0.03 &  \\ 
&  Graduate degree & 0.13 & 0.10 & 0.16 & *** &   & -0.04 & -0.17 & 0.09 &  \\ [3pt]
\bf Gender $(\gamma)$& Female & \multicolumn{5}{l}{0.00} & \multicolumn{4}{l}{\phantom{-}0.00} \\
&  Male & 0.08 & 0.06 & 0.10 & *** &   & 0.05 & -0.02 & 0.12 &  \\
\cline{3-6} \cline{8-11}
\bf Random Effects &  lineup $(\sigma_\ell)$& \multicolumn{5}{l}{0.29} & \multicolumn{4}{l}{\phantom{-}2.27} \\ 
                   &  Error $(\sigma)$& \multicolumn{5}{l}{0.69} & \multicolumn{4}{l}{\phantom{-}0.89}  \\ 
\cline{3-6} \cline{8-11}\\
\hline
\multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
\end{tabular}}
\end{table}

% % latex table generated in R 3.4.1 by xtable 1.8-2 package
% % Wed Oct 18 17:22:24 2017
% \begin{table}[ht]
% \centering
% \begin{tabular}{rrrrllrrrl}
%   \hline
%  & Est & LB & UB & pvalue & g1 & Est & LB & UB & pvalue \\ 
%   \hline
% $\mu$ & 3.16 & 3.12 & 3.20 & *** &   & -0.63 & -0.81 & -0.45 & *** \\ 
%   26-30 & 0.05 & 0.03 & 0.07 & *** &   & 0.03 & -0.06 & 0.13 &  \\ 
%   31-35 & 0.06 & 0.03 & 0.09 & *** &   & 0.12 & 0.02 & 0.23 & * \\ 
%   36-40 & 0.22 & 0.19 & 0.25 & *** &   & 0.29 & 0.17 & 0.41 & *** \\ 
%   41-45 & 0.17 & 0.13 & 0.21 & *** &   & 0.19 & 0.03 & 0.34 & * \\ 
%   46-50 & 0.28 & 0.23 & 0.32 & *** &   & 0.20 & 0.01 & 0.39 & * \\ 
%   above 50 & 0.34 & 0.31 & 0.38 & *** &   & 0.16 & 0.02 & 0.29 & * \\ 
%   India & 0.23 & 0.20 & 0.25 & *** &   & -0.19 & -0.28 & -0.10 & *** \\ 
%   Rest of the world & 0.14 & 0.11 & 0.17 & *** &   & 0.06 & -0.05 & 0.18 &  \\ 
%   Under grad courses & 0.13 & 0.09 & 0.16 & *** &   & 0.00 & -0.14 & 0.14 &  \\ 
%   Under grad degree & 0.14 & 0.10 & 0.17 & *** &   & 0.16 & 0.03 & 0.29 & * \\ 
%   Graduate courses & 0.02 & -0.02 & 0.06 &  &   & -0.13 & -0.29 & 0.03 &  \\ 
%   Graduate degree & 0.13 & 0.10 & 0.16 & *** &   & -0.04 & -0.17 & 0.09 &  \\ 
%   Male & 0.08 & 0.06 & 0.10 & *** &   & 0.05 & -0.02 & 0.12 &  \\ 
%   lineup & 0.29 &  &  &  &   & 2.27 &  &  &  \\ 
%    \hline
% \end{tabular}
% \end{table}

@@

While some of the demographic factors are strongly statistically significant, the main source of variation in detection rate is the lineup difficulty. For example, let's examine the effect of an undergraduate degree. To see just how large the effect is, we examine the change in detection rate for a (hypothetical) 18-25 year old female in the United States, with some graduate course work as compared to an undergraduate degree, for an average difficulty lineup (random effect of zero). Plugging in the relevant quantities to the fitted model gives a difference equal to:
$$ \frac {\exp(-0.63-0.13)}{1+\exp(-0.63-0.13)}-  \frac {\exp(-0.63+0.16)}{1+\exp(-0.63+0.16)} = 0.319- 0.385 =-0.066.$$ 
The person with some graduate course work picks the data plot  on average in 31.9\% of lineups of average difficulty, as compared to 38.5\% if they have an undergraduate degree. This difference is reduced to 4\% for a lineup with one standard deviation order of magnitude difference in difficulty. For two standard deviations it further reduces to 0.6\%. Thus, although there is a statistically significant difference in the proportion at which participants identify the  data plot for some demographic factors, these are not  significant in any practical sense.  Figure~\ref{practical_impact_und_und_graduate.pdf} illustrates this example showing fitted models for a US 18-25 female with either a high school education or an undergraduate degree.  Similar calculations show the same negligible impact of age level 36-40 (\Sexpr{round(exp(-0.63+0.29)/(1+ exp(-0.63+0.29)) - exp(-0.63)/(1 + exp(-0.63)),3)} at most)  and country India (\Sexpr{round(exp(-0.63-0.19)/(1+ exp(-0.63-0.19)) - exp(-0.63)/(1 + exp(-0.63)), 3)} at most) on the probability of a correct response. Thus even though some of the demographic factors are statistically significant, practically, demographics do not substantially influence the results.

\begin{figure}[htbp] 
   \centering
%   \includegraphics[width=6.5in]{learning_trend.pdf} 
    \includegraphics[width=3.2in]{practical_impact_und_graduate.pdf} 
    \includegraphics[width=3.2in]{practical_impact_degree.pdf} 
   \caption{Proportion of estimated data identifications by an 18-25 year old female in the United States with an under graduate degree compared to a high school degree. Even though having an under graduate degree leads to  statistically significant increase, the effect size in the dependent variable is  0.04 which is  negligible from a practical point of view. The difference diminishes as we move away one or two standard deviations ($\sigma_{\ell} = 2.27$) 
   of lineup variability.}
   \label{practical_impact_und_und_graduate.pdf}
\end{figure}

@@

% latex table generated in R 3.4.1 by xtable 1.8-2 package
% Wed Oct 18 23:54:25 2017
\begin{table}[hbtp]
\centering
\caption{\label{tbl:model_result_response}Parameter estimates of models \eqref{eqn:trend_response} fitted to data from three different experiments using detection rate as the response to assess learning trend. Attempt number is fitted as a factor to enable modeling any sort of non-linear learning trend. Only experiment~6 shows some evidence of a learning trend, with detection rate essentially increasing as attempts increase.}
\scalebox{.9}{
\begin{tabular}{rr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
\cline{2-5} \cline{7-10} \cline{12-15} 
& \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
\cline{2-5} \cline{7-10} \cline{12-15} 
Effect & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  & \hspace{1cm} & Est & 2.5\% & 97.5\% &  \\ 
\cline{2-5} \cline{7-10} \cline{12-15} 
Fixed \\
$\mu$ & -1.32 & -1.73 & -0.92 & *** &   & -0.16 & -0.46 & 0.13 &  &   & -1.73 & -2.81 & -0.66 & ** \\ [3pt]
  $\alpha2$ & 0.30 & -0.17 & 0.78 &  &   & 0.26 & -0.07 & 0.58 &  &   & -0.40 & -1.23 & 0.44 &  \\ 
  $\alpha3$ & -0.20 & -0.68 & 0.29 &  &   & 0.31 & -0.02 & 0.63 & . &   & -0.15 & -0.99 & 0.68 &  \\ 
  $\alpha4$ & 0.13 & -0.35 & 0.61 &  &   & 0.34 & 0.01 & 0.66 & * &   & -0.41 & -1.24 & 0.42 &  \\ 
  $\alpha5$ & 0.35 & -0.14 & 0.83 &  &   & 0.34 & 0.02 & 0.67 & * &   & -0.10 & -0.92 & 0.73 &  \\ 
  $\alpha6$ & 0.10 & -0.40 & 0.60 &  &   & 0.20 & -0.12 & 0.53 &  &   & 0.03 & -0.84 & 0.89 &  \\ 
  $\alpha7$ & 0.33 & -0.17 & 0.82 &  &   & 0.11 & -0.22 & 0.43 &  &   & -0.00 & -0.84 & 0.84 &  \\ 
  $\alpha8$ & -0.01 & -0.51 & 0.50 &  &   & 0.31 & -0.01 & 0.64 & . &   & -0.06 & -0.88 & 0.76 &  \\ 
  $\alpha9$ & -0.20 & -0.70 & 0.30 &  &   & 0.34 & 0.01 & 0.67 & * &   & 0.21 & -0.67 & 1.08 &  \\ 
  $\alpha10$ & 0.51 & 0.02 & 1.01 & * &   & 0.19 & -0.14 & 0.53 &  &   & -0.20 & -1.07 & 0.67 &  \\ 
  Random \\
  $\sigma^2_u$ & \multicolumn{5}{l}{\phantom{-}0.91}  & \multicolumn{5}{l}{\phantom{-}0.89} & 0.82   \\ 
  $\sigma^2_a$ & \multicolumn{5}{l}{\phantom{-}0.04}  & \multicolumn{5}{l}{\phantom{-}0.04} & 0.05   \\ 
  $\sigma^2_l$ & \multicolumn{5}{l}{\phantom{-}1.47}  & \multicolumn{5}{l}{\phantom{-}1.43} & 3.38   \\ \cline{2-5} \cline{7-10} \cline{12-15} 
\\
\hline
\multicolumn{10}{l}{Signif. codes:  *** $<$ 0.001 $\le$ ** $<$ 0.01 $\le$ * $<$ 0.05 $\le$ . $<$ 0.1 $\le$ ` ' $<$ 1}
\end{tabular}}
\end{table}

% 
% 
% % latex table generated in R 3.4.1 by xtable 1.8-2 package
% % Wed Oct 18 23:54:25 2017
% \begin{table}[ht]
% \centering
% \begin{tabular}{rrrrllrrrllrrrl}
%   \hline
%  & Est & LB & UB & pvalue & g1 & Est.1 & LB.1 & UB.1 & pvalue.1 & g2 & Est.2 & LB.2 & UB.2 & pvalue.2 \\ 
%   \hline
% $\mu$ & -1.32 & -1.73 & -0.92 & *** &   & -0.16 & -0.46 & 0.13 &  &   & -1.73 & -2.81 & -0.66 & ** \\ 
%   $\alpha2$ & 0.30 & -0.17 & 0.78 &  &   & 0.26 & -0.07 & 0.58 &  &   & -0.40 & -1.23 & 0.44 &  \\ 
%   $\alpha3$ & -0.20 & -0.68 & 0.29 &  &   & 0.31 & -0.02 & 0.63 & . &   & -0.15 & -0.99 & 0.68 &  \\ 
%   $\alpha4$ & 0.13 & -0.35 & 0.61 &  &   & 0.34 & 0.01 & 0.66 & * &   & -0.41 & -1.24 & 0.42 &  \\ 
%   $\alpha5$ & 0.35 & -0.14 & 0.83 &  &   & 0.34 & 0.02 & 0.67 & * &   & -0.10 & -0.92 & 0.73 &  \\ 
%   $\alpha6$ & 0.10 & -0.40 & 0.60 &  &   & 0.20 & -0.12 & 0.53 &  &   & 0.03 & -0.84 & 0.89 &  \\ 
%   $\alpha7$ & 0.33 & -0.17 & 0.82 &  &   & 0.11 & -0.22 & 0.43 &  &   & -0.00 & -0.84 & 0.84 &  \\ 
%   $\alpha8$ & -0.01 & -0.51 & 0.50 &  &   & 0.31 & -0.01 & 0.64 & . &   & -0.06 & -0.88 & 0.76 &  \\ 
%   $\alpha9$ & -0.20 & -0.70 & 0.30 &  &   & 0.34 & 0.01 & 0.67 & * &   & 0.21 & -0.67 & 1.08 &  \\ 
%   $\alpha10$ & 0.51 & 0.02 & 1.01 & * &   & 0.19 & -0.14 & 0.53 &  &   & -0.20 & -1.07 & 0.67 &  \\ 
%   $\sigma^2_u$ & 0.91 &  &  &  &   & 0.89 &  &  &  &   & 0.82 &  &  &  \\ 
%   $\sigma^2_a$ & 0.04 &  &  &  &   & 0.04 &  &  &  &   & 0.05 &  &  &  \\ 
%   $\sigma^2_l$ & 1.47 &  &  &  &   & 1.43 &  &  &  &   & 3.38 &  &  &  \\ 
%    \hline
% \end{tabular}
% \end{table}
% 
%

@@

% latex table generated in R 3.4.1 by xtable 1.8-2 package
% Wed Oct 18 23:37:55 2017
\begin{table}[bhtp]
\caption{\label{tbl:model_result_time}Parameter estimates of model~\eqref{eqn:trend_time} fitted for log time taken to evaluate a lineup. Both fixed effects parameters of Attempt ($\alpha_1$ and $\alpha$) are highly significant for all three experiments~5,~6 and~7.}
\centering
\scalebox{0.9}{
\begin{tabular}{lr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{( }r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}cr>{(}r<{,}>{\hspace{-.1in}}l<{)}>{\hspace{-.15in}}l<{}}
\cline{2-5} \cline{7-10} \cline{12-15} 
& \multicolumn{4}{c} {Experiment~5} & &\multicolumn{4}{c} {Experiment~6} && \multicolumn{4}{c} {Experiment~7}\\
